---
title: "Conditional Formatting with Color Tiles"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(dplyr)
library(palmerpenguins)
```

<i>Please note: the `color_tiles()` formatter used in the examples below are currently only available in the development version of `{reactablefmtr}` and not on CRAN as of this moment. Please download the development version of the package from GitHub via `remotes::install_github("kcuilla/reactablefmtr")` if would like to use `color_tiles()`.</i>

## Example

Let's start with an example using the palmerpenguins [dataset](https://allisonhorst.github.io/palmerpenguins/) from Allison Horst, Alison Hill, and Kristen Gorman.   

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, bill_length_mm, bill_depth_mm, flipper_length_mm)

reactable(data)
```

An alternative method to `color_scales()` is the `color_tiles()` formatter. The main difference between `color_scales()` and `color_tiles()` is that the `color_tiles()` formatter provides:

- Round color tiles around values vs coloring the entire background of the cell
- Called within the `cell` argument of reactable vs the `style` argument
- Since it is called within the `cell` argument, you can format numbers directly using the scales package

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(align = "center", # center column header
                                    cell = color_tiles(data))))
```

<br>

## Custom Color Palettes

If we want to show a different color palette than the default, we can call them within the `colors` argument just like we would with `color_scales()`: 

```{r}
my_color_pal = c("purple", "white", "green")

reactable(data,
          defaultColDef = colDef(cell = color_tiles(data, colors = my_color_pal)))
```

And just like with `color_scales()`, you are now free to use dark color palettes:

```{r}
library(viridis)
reactable(data,
          defaultColDef = colDef(cell = color_tiles(data, colors = viridis::magma(5))))
```

<br>

## Formatting Numbers

You can now use the scales package to format numbers within `color_tiles()` similarly to how you would within ggplot2.

For example, if we wanted to round the numbers in the `bill_length_mm` column to the nearest whole number and add commas to the `body_mass_g` column, we could do so using the `number_fmt` argument below and calling the comma formatter directly from the scales package:

```{r}
df <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, bill_length_mm, body_mass_g)

reactable(df,
          defaultColDef = colDef(align = "center", # center column header
                                 cell = color_tiles(df, 
                                                    number_fmt = scales::comma)))
```

## Using Span

Just like with `color_scales()`, by default, `color_tiles()` conditionally assigns colors to values based on their relation to other values within that particular column. However, if the table you're showing is row-wise data, such as average temperatures by month for each year, then it will be difficult to compare how temperatures compare in each month as shown below:

<i>Note: the dataset for this example is sourced from the reactable [demo cookbook](https://glin.github.io/reactable/articles/cookbook/cookbook.html#conditional-formatting-1)</i>

```{r}
dimnames <- list(start(nottem)[1]:end(nottem)[1], month.abb)
temps <- matrix(nottem, ncol = 12, byrow = TRUE, dimnames = dimnames)
temps <- as_tibble(temps, rownames = "Year")

reactable(
  temps,
  defaultColDef = colDef(
    cell = color_tiles(temps, colors = c("#1e90ff", "#ffffff", "#ff3030")),
    minWidth = 50))
```

By including `span = TRUE` within our `color_tiles()` formatter, we can conditionally assign colors to the values based on their relation to other values within the entire dataset, instead of within each column. Now our table displaying temperatures is much easier to read when comparing temperatures across months:

```{r}
reactable(
  temps,
  defaultColDef = colDef(
    cell = color_tiles(temps, span = TRUE, colors = c("#1e90ff", "#ffffff", "#ff3030")),
    minWidth = 50))
```

<i>Note: this feature is only available in the development version and is not on CRAN at the moment.</i>
