---
title: "Conditional Formatting with Color Scales"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(dplyr)
library(tibble)
library(palmerpenguins)
```

## Example

Let's start with an example using the palmerpenguins [dataset](https://allisonhorst.github.io/palmerpenguins/) from Allison Horst, Alison Hill, and Kristen Gorman.   

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, bill_length_mm, bill_depth_mm, flipper_length_mm)

reactable(data)
```

If we want to add color scales to the `bill_length_mm` column, all we need to do is add `color_scales(data)` within the `style` argument of `colDef`, and a normalized red-white-blue color scale will be applied:

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data))))
```

You may recall the original method of doing this via `{reactable}` looks like this:

```{r}
color_pal <- function(x) rgb(colorRamp(c("#ff3030", "#ffffff", "#1e90ff"))(x), maxColorValue = 255)

reactable(data, columns = list(
  bill_length_mm = colDef(style = function(value) {
    normalized <- (value - min(data$bill_length_mm, na.rm = TRUE)) / (max(data$bill_length_mm, na.rm = TRUE) - min(data$bill_length_mm, na.rm = TRUE))
    color <- color_pal(normalized)
    list(background = color)
  })
))
```

As you can see, `color_scales()` significantly reduces the amount of code needed, is more reproducible, and is easier to customize as I'm about to further show below.

<br>

## Custom Color Palettes

If we want to show a different color palette than the default, we can call them within the `colors` argument like so: 

```{r}
my_color_pal = c("purple", "lightgrey", "darkgreen")

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = my_color_pal)),
            bill_depth_mm = colDef(style = color_scales(data, colors = my_color_pal)),
            flipper_length_mm = colDef(style = color_scales(data, colors = my_color_pal))))
```

The order of the color palette matters. In the example above, the color purple is assigned to the lowest values, lightgrey is assigned to middle values, and green is assigned to highest values. If we want green to represent the lowest values and purple to represent the highest values, we can just reverse the order of the color palette:

```{r}
my_color_pal = c("darkgreen", "lightgrey", "purple")

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = my_color_pal)),
            bill_depth_mm = colDef(style = color_scales(data, colors = my_color_pal)),
            flipper_length_mm = colDef(style = color_scales(data, colors = my_color_pal))))
```

In the examples above, we've only used three colors in each color palette, but we can use as many colors as we desire:

```{r}
my_color_pal = c("red", "orange", "yellow", "green", "skyblue", "purple")

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = my_color_pal))))
```

We may also use color palettes from other packages, such as the "Chevalier1" color palette from the `{wesanderson}` [package](https://github.com/karthik/wesanderson):

```{r, warning=FALSE}
library(wesanderson)

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = wes_palette("Chevalier1"))),
            bill_depth_mm = colDef(style = color_scales(data, colors = wes_palette("Chevalier1"))),
            flipper_length_mm = colDef(style = color_scales(data, colors = wes_palette("Chevalier1")))))
```

Previously in {reactable}, using dark color palettes such as the "magma" color set from {viridis} was troublesome since you couldn't see the values in the cells with dark backgrounds. 

But, with {reactablefmtr}, the colors of the values automatically are changed to white if the colors are dark. This option is already turned on by default, but if you don't want to use this feature, you can turn it off by setting `bright_values = FALSE`.

```{r}
library(viridis)

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = viridis::magma(5))),
            bill_depth_mm = colDef(style = color_scales(data, colors = viridis::magma(5))),
            flipper_length_mm = colDef(style = color_scales(data, colors = viridis::magma(5)))))
```

## Using Span

By default, `color_scales()` conditionally assigns colors to values based on their relation to other values within that particular column. However, if the table you're showing is row-wise data, such as average temperatures by month for each year, then it will be difficult to compare how temperatures compare in each month as shown below:

<i>Note: the dataset for this example is sourced from the reactable [demo cookbook](https://glin.github.io/reactable/articles/cookbook/cookbook.html#conditional-formatting-1)</i>

```{r}
dimnames <- list(start(nottem)[1]:end(nottem)[1], month.abb)
temps <- matrix(nottem, ncol = 12, byrow = TRUE, dimnames = dimnames)
temps <- as_tibble(temps, rownames = "Year")

reactable(
  temps,
  defaultColDef = colDef(
    style = color_scales(temps, colors = c("#1e90ff", "#ffffff", "#ff3030")),
    minWidth = 50))
```

By including `span = TRUE` within our `color_scales()` formatter, we can conditionally assign colors to the values based on their relation to other values within the entire dataset, instead of within each column. Now our table displaying temperatures is much easier to read when comparing temperatures across months:

```{r}
reactable(
  temps,
  defaultColDef = colDef(
    style = color_scales(temps, span = TRUE, colors = c("#1e90ff", "#ffffff", "#ff3030")),
    minWidth = 50))
```

