---
title: "Conditional Formatting with Color Scales"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
library(reactable)
library(reactablefmtr)
library(dplyr)
library(palmerpenguins)
```

## Example

Let's start with an example using the palmerpenguins [dataset](https://allisonhorst.github.io/palmerpenguins/) from Allison Horst, Alison Hill, and Kristen Gorman.   

```{r, echo=TRUE}
data <- sample_n(penguins,50) %>% select(species, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

reactable(data)
```

If we want to add color scales to the `bill_length_mm` column, all we need to do is add `color_scales(data)` within the `style` argument of `colDef`, and a normalized red-white-blue color scale will be applied:

```{r, echo=TRUE}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data))))
```

You may recall the original method of doing this via `{reactable}` looks like this:

```{r, echo=TRUE}
color_pal <- function(x) rgb(colorRamp(c("#ff3030", "#ffffff", "#1e90ff"))(x), maxColorValue = 255)

reactable(data, columns = list(
  bill_length_mm = colDef(style = function(value) {
    normalized <- (value - min(data$bill_length_mm, na.rm = TRUE)) / (max(data$bill_length_mm, na.rm = TRUE) - min(data$bill_length_mm, na.rm = TRUE))
    color <- color_pal(normalized)
    list(background = color)
  })
))
```

As you can see, `color_scales()` significantly reduces the amount of code needed, is more reproducible, and is easier to customize as I'm about to further show below.

## Custom Color Palettes

If we want to show a different color palette than the default, we can call them within the `colors` argument like so: 

```{r, echo=TRUE}
my_color_pal = c("purple", "lightgrey", "green")

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = my_color_pal))))
```

The order of the color palette matters. In the example above, the color purple is assigned to the lowest values, lightgrey is assigned to middle values, and green is assigned to highest values. If we want green to represent the lowest values and purple to represent the highest values, we can just reverse the order of the color palette:

```{r, echo=TRUE}
my_color_pal = c("green", "lightgrey", "purple")

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = my_color_pal))))
```

In the examples above, we've only used three colors in each color palette, but we can use as many colors as we desire:

```{r, echo=TRUE}
my_color_pal = c("red", "orange", "yellow", "green", "skyblue", "purple")

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = my_color_pal))))
```

We may also use color palettes from other packages, such as the "Chevalier1" color palette from the `{wesanderson}` [package](https://github.com/karthik/wesanderson):

```{r, echo=TRUE, warning=FALSE}
library(wesanderson)

reactable(data,
          columns = list(
            bill_length_mm = colDef(style = color_scales(data, colors = wes_palette("Chevalier1")))))
```

## Apply to All Columns

Lastly, if we want to apply `color_scales()` to every numeric column, we can simply use `defaultColDef` as shown below:

```{r, echo=TRUE}
reactable(data,
          defaultColDef = colDef(style = color_scales(data)))
```

