---
title: "Getting Started with Color Tiles"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(tidyverse)
library(gapminder)
```

## Default

With `{reactablefmtr}`, we can easily apply color tiles to columns within a table by using `color_tiles()` within `cell` of `reactable::colDef()`.

By default a normalized orange-white-blue color scale is applied to each column.

```{r}
data <- MASS::Cars93 %>% 
  filter(Type %in% c("Compact", "Sporty", "Van")) %>% 
  select(c("Make", "Type", "Horsepower", "MPG.city", "MPG.highway")) %>% 
  tail(10)

reactable(
  data,
  defaultColDef = colDef(
    cell = color_tiles(data)
  )
)
```


## Custom Color Palettes

If we want to show a different color palette than the default, we can call them within the `colors` argument like so: 

```{r}
my_color_pal = c("#e5f5e0", "#a1d99b", "#31a354")

reactable(
  data,
  defaultColDef = colDef(
    cell = color_tiles(data, colors = my_color_pal)
  )
)
```

The opacity of the colors can be controlled by assigning a value between 0 (fully transparent) and 1 (fully opaque).

```{r}
reactable(
  data,
  defaultColDef = colDef(
    cell = color_tiles(data, colors = my_color_pal, opacity = 0.5)
  )
)
```


The order of the color palette matters. In the example above, the color purple is assigned to the lowest values, lightgrey is assigned to middle values, and green is assigned to highest values. 

We may also use color palettes from other packages, such as the "Mako" color palette from the `{viridis}` package:

```{r}
library(viridis)

reactable(
  data,
  defaultColDef = colDef(
    cell = color_tiles(data, colors = viridis::mako(5))
  )
)
```


## Formatting Numbers

Numbers can be formatted within the `number_fmt` argument in `color_tiles()`. One method of formatting the numbers is by utilizing the formatters from the `{scales}` package, and the numbers can be formatted in the same way as they are in `{ggplot2}`.

```{r}
reactable(
  data,
  columns = list(
    Horsepower = colDef(cell = color_tiles(data, number_fmt = scales::number_format(suffix = " hp"))),
    MPG.city = colDef(cell = color_tiles(data, number_fmt = scales::number_format(suffix = " mpg"))),
    MPG.highway = colDef(cell = color_tiles(data, number_fmt = scales::number_format(suffix = " mpg")))
  )
)
```

## Adding a Box Shadow

Box shadows can be added to the tiles to create a "3-D" effect via `box_shadow`.

```{r}
reactable(
  data,
  defaultColDef = colDef(
    cell = color_tiles(data,
                       box_shadow = TRUE)
  )
)
```


## Assigning Colors from Another Column

Colors can be conditionally assigned to values based on another column by using `color_ref`.

In the example below, we assigned a blue color to Compact cars, a red color to Sporty cars, and a gold color to Vans using `dplyr::case_when()`.

Then within `color_tiles()`, we reference the name of the conditional column we just created to apply the colors to the values in MPG.city and MPG.highway.

```{r}
car_data <- data %>% 
  mutate(car_colors = dplyr::case_when(
  Type == "Compact" ~ "dodgerblue",
  Type == "Sporty" ~ "tomato",
  Type == "Van" ~ "gold",
  TRUE ~ "other"
  ))

reactable(
  car_data,
  columns = list(
    Make = colDef(minWidth = 150),
    car_colors = colDef(show = FALSE),
    Horsepower = colDef(cell = color_tiles(car_data, color_ref = "car_colors")),
    MPG.city = colDef(cell = color_tiles(car_data, color_ref = "car_colors")),
    MPG.highway = colDef(cell = color_tiles(car_data, color_ref = "car_colors"))
  )
)
```

We can further apply the conditional colors to the entire dataset by setting the style within `defaultColDef`:

```{r}
reactable(
  car_data,
    defaultColDef = colDef(
    cell = color_tiles(car_data, color_ref = "car_colors")
  ),
  columns = list(
    Make = colDef(minWidth = 150),
    car_colors = colDef(show = FALSE)
  )
)
```


The same conditional coloring can be applied based on numeric conditions as well. For example, if we wanted to highlight which cars have an MPG.city value of 23 or greater, we could use the same method as above but apply the conditions based on the MPG.city column instead of the Type column.  


```{r}
car_data <- car_data %>% 
  mutate(mpg_colors = dplyr::case_when(
  MPG.city >= 23 ~ "darkgreen",
  TRUE ~ "grey"
  ))

reactable(
  car_data,
  defaultSorted = "MPG.city",
  defaultSortOrder = "desc",
  defaultColDef = colDef(
    cell =color_tiles(car_data, color_ref = "mpg_colors")
    ),
  columns = list(
    Make = colDef(minWidth = 150),
    car_colors = colDef(show = FALSE),
    mpg_colors = colDef(show = FALSE))
)
```



## Row-wise Styling

By default, `color_tiles()` conditionally assigns colors to values based on their relation to other values within that particular column. However, if the table you're showing is row-wise data, such as average temperatures by month for each year, then it will be difficult to compare how temperatures compare in each month:

By including `span = TRUE` within our `color_tiles()` formatter, we can conditionally assign colors to the values based on their relation to other values within the entire dataset, instead of within each column. Now our table displaying temperatures is much easier to read when comparing temperatures across months:

<i>Note: the dataset for this example is sourced from the reactable [demo cookbook](https://glin.github.io/reactable/articles/cookbook/cookbook.html#conditional-formatting-1)</i>

```{r}
dimnames <- list(start(nottem)[1]:end(nottem)[1], month.abb)
temps <- matrix(nottem, ncol = 12, byrow = TRUE, dimnames = dimnames)
temps <- as_tibble(temps, rownames = "Year")
temppal <- c("#36a1d6", "#76b8de", "#a0bfd9", "#ffffff", "#d88359", "#d65440", "#c62c34")

reactable(
  temps,
  defaultColDef = colDef(
    cell =color_tiles(temps, span = TRUE, colors = temppal),
    minWidth = 50
    )
)
```


If we only wanted to apply color tiles to some of the months, we can do so by referencing either the numeric positions of the columns or the column names within `span`:

```{r}
reactable(
  temps,
  defaultColDef = colDef(
    cell =color_tiles(temps, span = 4:7, colors = temppal),
    minWidth = 50))
```


Lastly, if you wanted to completely hide the text, you could do this by setting `show_text = FALSE`, which displays the table as a heatmap as shown below:

```{r}
population_data <- gapminder %>% 
  filter(continent == "Americas") %>%
  mutate(country = as.character(country),
         year = paste0("'", str_sub(year, 3, 4))) %>% 
  select(country, year, lifeExp) %>%
  pivot_wider(names_from = year, values_from = lifeExp) 

reactable(
  population_data,
  pagination = FALSE,
  showSortIcon = FALSE,
  defaultSorted = "'52",
  defaultSortOrder = "desc",
  defaultColDef = colDef(
    maxWidth = 50,
    cell = color_tiles(population_data, show_text = FALSE, span = TRUE)
  ),
  columns = list(
    country = colDef(
      maxWidth = 160
    )
  )
) %>% 
  add_title("Average Life Expectancy") %>% 
  add_source("Data sourced from the {gapminder} package") 
```

