---
title: "Conditional Formatting with Color Tiles"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(dplyr)
library(palmerpenguins)
```

<i>Please note: the `color_tiles()` formatter used in the examples below are currently only available in the development version of `{reactablefmtr}` and not on CRAN as of this moment. Please download the development version of the package from GitHub via `remotes::install_github("kcuilla/reactablefmtr")` if would like to use `color_tiles()`.</i>

## Example

Let's start with an example using the palmerpenguins [dataset](https://allisonhorst.github.io/palmerpenguins/) from Allison Horst, Alison Hill, and Kristen Gorman.   

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, bill_length_mm, bill_depth_mm, flipper_length_mm)

reactable(data)
```

An alternative method to `color_scales()` is the `color_tiles()` formatter. The main difference between `color_scales()` and `color_tiles()` is that the `color_tiles()` formatter provides:

- Round color tiles around values vs coloring the entire background of the cell
- Called within the `cell` argument of reactable vs the `style` argument
- Since it is called within the `cell` argument, you can format numbers directly using the scales package

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(align = "center", # center column header
                                    cell = color_tiles(data))))
```

<br>

## Custom Color Palettes

If we want to show a different color palette than the default, we can call them within the `colors` argument just like we would with `color_scales()`: 

```{r}
my_color_pal = c("purple", "white", "green")

reactable(data,
          defaultColDef = colDef(cell = color_tiles(data, colors = my_color_pal)))
```

And just like with `color_scales()`, you are now free to use dark color palettes:

```{r}
library(viridis)
reactable(data,
          defaultColDef = colDef(cell = color_tiles(data, colors = viridis::magma(5))))
```

<br>

## Formatting Numbers

You can now use the scales package to format numbers within `color_tiles()` similarly to how you would within ggplot2.

For example, if we wanted to round the numbers in the `bill_length_mm` column to the nearest whole number and add commas to the `body_mass_g` column, we could do so using the `number_fmt` argument below and calling the comma formatter directly from the scales package:

```{r}
df <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, bill_length_mm, body_mass_g)

reactable(df,
          defaultColDef = colDef(align = "center", # center column header
                                 cell = color_tiles(df, 
                                                    number_fmt = scales::comma)))
```


