---
title: "Conditional Formatting with Data Bars"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(htmltools)
library(dplyr)
library(palmerpenguins)
```

<i>Please note: the animation of the data bars, `data_bars_gradient()` formatter, and the `number_fmt` argument used in the examples below are currently only available in the development version of `{reactablefmtr}` and not on CRAN as of this moment. Please download the development version of the package from GitHub via `remotes::install_github("kcuilla/reactablefmtr")` if would like to use those features.</i>

## Example 1 - data_bars()

The `data_bars()` function can be used to apply conditional data bars to <b>positive</b> values.

For example, let's say that we want to add conditional data bars to the `body_mass_g` column of the palmerpenguins [dataset](https://allisonhorst.github.io/palmerpenguins/).

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, island, body_mass_g)

reactable(data)
```

To do this, we just need to add `data_bars(data)` within the `cell` argument of `colDef`, and a default blue data bar will be applied to the `body_mass_g` column:

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars(data))))
```

Adding conditional data bars to a column is significantly easier compared to the original method via `{reactable}` (and you also lose out on the animation of the bars too):

```{r}
bar_chart <- function(label, width = "100%", height = "16px", fill = "#1e90ff", background = NULL) {
  bar <- div(style = list(background = fill, width = width, height = height))
  chart <- div(style = list(flexGrow = 1, marginLeft = "8px", background = background), bar)
  div(style = list(display = "flex", alignItems = "center"), label, chart)
}

reactable(data, columns = list(
  body_mass_g = colDef(cell = function(value) {
    width <- paste0(value / max(data$body_mass_g, na.rm = TRUE) * 100, "%")
    bar_chart(value, width = width)
  })
))
```

<br>

## Custom Color Palettes

By default, the data bars are blue and the background is white. Both of these colors can be modified as shown below:

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars(data, colors = "red", background = "lightgrey"))))
```

You may be wondering why the `colors` argument is plural and not singular. That is because you can apply multiple colors to the data bars, which is something that we could not previously do using the original method from `{reactable}` shown in Example 1 above.

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars(data, colors = c("#ff3030", "#ffffff", "#1e90ff"), background = "lightgrey"))))
```

<br>

## Apply to All Columns

We can apply data bars across all columns using `defaultColDef`: 

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, island, bill_length_mm, body_mass_g)

reactable(data,
          defaultColDef = colDef(cell = data_bars(data)))
```

<br>

## Aligning Values

You may have noticed that when we applied the data bars across all columns, the `bill_length_mm` column did not fully align properly due to the different value lengths. We can correct this by using a monospaced font as shown in the [Building the Twitter Follows Demo](https://glin.github.io/reactable/articles/building-twitter-followers.html#add-bar-charts-1) from the `{reactable}` package documentation:

```{r}
reactable(data,
          defaultColDef = colDef(cell = data_bars(data),
                                 style = list(fontFamily = "monospace", whiteSpace = "pre")))
```

Now all the values are aligned properly with the data bars!

If we would like to format the numbers, such as rounding the numbers in the `bill_length_mm` to the nearest whole number and adding commas to the `body_mass_g` column, we can do so calling the comma formatter from the scales package within `number_fmt`:

```{r}
reactable(data,
          defaultColDef = colDef(cell = data_bars(data,
                                                  number_fmt = scales::comma),
                                 style = list(fontFamily = "monospace", whiteSpace = "pre")))
```


<hr>


## Example 2 - data_bars_gradient()

If you would like to customize your `data_bars()` a step further and apply a linear color gradient, you can now use `data_bars_gradient()`.

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, island, body_mass_g)

reactable(data)
```

By default, a light blue-to-dark blue linear gradient is applied when applying `data_bars_gradient()` to a column:

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars_gradient(data))))
```

<br>

## Custom Color Palettes

You can provide as many colors as you want to your color gradient as long as you provide at least two colors:

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars_gradient(data, 
                                                           colors = c("firebrick", "red", "pink")))))
```

If you always wanted to apply a rainbow gradient to your column, you can now do that!

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars_gradient(data, 
                                                           colors = c("red", "orange", "yellow", "green", "blue", "indigo", "violet")))))
```

Just like with `data_bars()`, you can also apply a singular background color:

```{r}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars_gradient(data,
                                                           colors = c("firebrick", "red", "pink"),
                                                           background = "grey"))))
```

<br>

## Apply to All Columns

We can apply `data_bars_gradient()` across all columns using `defaultColDef`: 

```{r}
data <- sample_n(penguins,50) %>% # sample 50 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, island, bill_length_mm, body_mass_g)

reactable(data,
          defaultColDef = colDef(cell = data_bars_gradient(data)))
```

<br>

## Aligning Values

If your values contain different numeric lengths, you can use the same method of aligning them with the bars as you would with `data_bars()` as shown below:

```{r}
reactable(data,
          defaultColDef = colDef(cell = data_bars_gradient(data),
                                 style = list(fontFamily = "monospace", whiteSpace = "pre")))
```

You can also still use the comma formatter from the scales package to display commas in the values of any column containing values in the thousands or greater.

```{r}
reactable(data,
          defaultColDef = colDef(cell = data_bars_gradient(data,
                                                           number_fmt = scales::comma),
                                 style = list(fontFamily = "monospace", whiteSpace = "pre")))
```


<hr>


## Example 3 - data_bars_pos_neg()

If your dataset contains <b>negative</b> values, you can use `data_bars_pos_neg()` to display conditional data bars. 

For example, let's say we wanted to show the percent difference between the price of each car and the average price of all cars in the `Cars93` dataset from the `MASS` package:

```{r}
data <- sample_n(MASS::Cars93,15) %>% 
  select(Make, Price)

data <- data %>% 
  mutate("Price vs Avg" = round(Price/mean(Price)-1,2)*100)

reactable(data)
```

To to show data bars for both positive and negative values, we can use `data_bars_pos_neg()`:

```{r}
reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data))))
```

As with `data_bars()`, using `data_bars_pos_neg()` requires significantly less code vs the original method via `{reactable}` shown below (and again, you also lose out on the animation of the bars as well):

```{r}
bar_chart_pos_neg <- function(label, value, max_value = 1, height = "16px",
                              pos_fill = "green", neg_fill = "red") {
  neg_chart <- div(style = list(flex = "1 1 0"))
  pos_chart <- div(style = list(flex = "1 1 0"))
  width <- paste0(abs(value) / max(abs(data$`Price vs Avg`), na.rm = TRUE) * 100, "%")

  if (value < 0) {
    bar <- div(style = list(marginLeft = "8px", background = neg_fill, width = width, height = height))
    chart <- div(style = list(display = "flex", alignItems = "center", justifyContent = "flex-end"), label, bar)
    neg_chart <- tagAppendChild(neg_chart, chart)
  } else {
    bar <- div(style = list(marginRight = "8px", background = pos_fill, width = width, height = height))
    chart <- div(style = list(display = "flex", alignItems = "center"), bar, label)
    pos_chart <- tagAppendChild(pos_chart, chart)
  }

  div(style = list(display = "flex"), neg_chart, pos_chart)
}

reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = function(value) {
      label <- value
      bar_chart_pos_neg(label, value)})
))
```

<br>

## Custom Color Palettes

Changing the colors assigned to the negative and positive values can be done using the `colors` argument below:

```{r}
reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data, colors = c("orange", "dodgerblue")))))
```

Just like with `data_bars()` above, if we provide more than two colors, `data_bars_pos_neg()` will assign a normalized color palette to the column:

```{r}
reactable(data,
          defaultSortOrder = "desc", # sort by descending order
          defaultSorted = "Price vs Avg", # sort by 'Price vs Avg'
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data, colors = c("#1e90ff", "#ffd700", "#ff3030")))))
```

<br>

## Adding Percentages to Labels

If you recall in the example above, we calculated the `Price vs Avg` column as the percent difference between the car's price and the average for all cars in the dataset. If we want to display the percentages on the values next to the data bars, we can do so by calling the percent formatter from the scales package within `number_fmt`:

```{r, echo=TRUE}
data <- data %>% 
  mutate("Price vs Avg" = Price/mean(Price)-1)

reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data,
                                                             number_fmt = scales::percent))))
```

<br>

## Combining Multiple Formatters 

If we wanted to apply `data_bars()` to one column and `data_bars_pos_neg()` to another, it can be done as shown below:

```{r}
reactable(data,
          defaultColDef = colDef(style = list(fontFamily = "monospace", whiteSpace = "pre")), # apply monospace font across all columns
          columns = list(
            Price = colDef(align = "left", # align column header to the left
                           cell = data_bars(data)),
            `Price vs Avg` = colDef(align = "center", # align column header to the center
                                    cell = data_bars_pos_neg(data, 
                                                             number_fmt = scales::percent))))
```

Similarly, we could also use `data_bars_gradient()` instead of `data_bars()`:

```{r}
reactable(data,
          defaultColDef = colDef(style = list(fontFamily = "monospace", whiteSpace = "pre")), # apply monospace font across all columns
          columns = list(
            Price = colDef(align = "left", # align column header to the left
                           cell = data_bars_gradient(data)),
            `Price vs Avg` = colDef(align = "center", # align column header to the center
                                    cell = data_bars_pos_neg(data, 
                                                             colors = c("#1e90ff", "#ffd700", "#ff3030"), 
                                                             number_fmt = scales::percent))))
```

