---
title: "Conditional Formatting with Data Bars"
output:
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
---

```{r setup, include=FALSE}
library(reactable)
library(reactablefmtr)
library(htmltools)
library(dplyr)
library(palmerpenguins)
```

## Example 1 - Positive Values

Let's say that we want to add conditional data bars to the `body_mass_g` column of the palmerpenguins [dataset](https://allisonhorst.github.io/palmerpenguins/).

```{r, echo=TRUE}
data <- sample_n(penguins,50) %>% select(species, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

reactable(data)
```

To do this, we just need to add `data_bars(data)` within the `cell` argument of `colDef`, and a default blue data bar will be applied to the `body_mass_g` column:

```{r, echo=TRUE}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars(data))))
```

Adding conditional data bars to a column is significantly easier compared to the original method via `{reactable}`:

```{r, echo=TRUE}
bar_chart <- function(label, width = "100%", height = "16px", fill = "#1e90ff", background = NULL) {
  bar <- div(style = list(background = fill, width = width, height = height))
  chart <- div(style = list(flexGrow = 1, marginLeft = "8px", background = background), bar)
  div(style = list(display = "flex", alignItems = "center"), label, chart)
}

reactable(data, columns = list(
  body_mass_g = colDef(cell = function(value) {
    width <- paste0(value / max(data$body_mass_g, na.rm = TRUE) * 100, "%")
    bar_chart(value, width = width)
  })
))
```

## Custom Color Palettes

By default, the data bars are blue and the background is white. Both of these colors can be modified as shown below:

```{r, echo=TRUE}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars(data, colors = "red", background = "lightgrey"))))
```

You may be wondering why the `colors` argument is plural and not singular. That is because you can apply multiple colors to the data bars, which is something that we could not previously do using the original method from `{reactable}` shown in Example 1 above.

```{r, echo=TRUE}
reactable(data,
          columns = list(
            body_mass_g = colDef(cell = data_bars(data, colors = c("#ff3030", "#ffffff", "#1e90ff"), background = "lightgrey"))))
```

## Apply to All Columns

We can apply data bars across all columns using `defaultColDef`: 

```{r, echo=TRUE}
reactable(data,
          defaultColDef = colDef(cell = data_bars(data)))
```

## Aligning Values

You may have noticed that when we applied the data bars across all columns, the `bill_length_mm` and `bill_depth_mm` did not fully align properly due to the different value lengths. We can correct this by using a monospaced font as shown in the [Building the Twitter Follows Demo](https://glin.github.io/reactable/articles/building-twitter-followers.html#add-bar-charts-1) from the `{reactable}` package documentation:

```{r, echo=TRUE}
reactable(data,
          defaultColDef = colDef(cell = data_bars(data),
                                 style = list(fontFamily = "monospace", whiteSpace = "pre")))
```

Now all the values are aligned properly with the data bars!

## Example 2 - Negative Values

If your dataset contains negative values, you'll need to use `data_bars_pos_neg()`. 

For example, let's say we wanted to show the percent difference between the price of each car and the average price of all cars in the `Cars93` dataset from the `MASS` package:

```{r, echo=TRUE}
data <- MASS::Cars93[c("Make", "Price")]

data <- data %>% mutate("Price vs Avg" = round(Price/mean(Price)-1,2)*100)

reactable(data)
```

If we try displaying the `MPG.Difference` column with `data_bars()`, the negative values show as positive values:

```{r, echo=TRUE}
reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars(data))))
```

To fix this, we just need to replace `data_bars()` with `data_bars_pos_neg()`:

```{r, echo=TRUE}
reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data))))
```

As with `data_bars()`, using `data_bars_pos_neg()` requires significantly less code vs the original method via `{reactable}` shown below:

```{r, echo=TRUE}
bar_chart_pos_neg <- function(label, value, max_value = 1, height = "16px",
                              pos_fill = "green", neg_fill = "red") {
  neg_chart <- div(style = list(flex = "1 1 0"))
  pos_chart <- div(style = list(flex = "1 1 0"))
  width <- paste0(abs(value) / max(abs(data$`Price vs Avg`), na.rm = TRUE) * 100, "%")

  if (value < 0) {
    bar <- div(style = list(marginLeft = "8px", background = neg_fill, width = width, height = height))
    chart <- div(style = list(display = "flex", alignItems = "center", justifyContent = "flex-end"), label, bar)
    neg_chart <- tagAppendChild(neg_chart, chart)
  } else {
    bar <- div(style = list(marginRight = "8px", background = pos_fill, width = width, height = height))
    chart <- div(style = list(display = "flex", alignItems = "center"), bar, label)
    pos_chart <- tagAppendChild(pos_chart, chart)
  }

  div(style = list(display = "flex"), neg_chart, pos_chart)
}

reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = function(value) {
      label <- value
      bar_chart_pos_neg(label, value)})
))
```

## Custom Color Palettes

Changing the colors assigned to the negative and positive values can be done using the `colors` argument below:

```{r, echo=TRUE}
reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data, colors = c("orange", "dodgerblue")))))
```

Just like with `data_bars()` above, if we provide more than two colors, `data_bars_pos_neg()` will assign a normalized color palette to the column:

```{r, echo=TRUE}
reactable(data,
          defaultSortOrder = "desc", # sort by descending order
          defaultSorted = "Price vs Avg", # sort by 'Price vs Avg'
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data, colors = c("#1e90ff", "#ffd700", "#ff3030")))))
```

## Adding Percentages to Labels

If you recall in the example above, we calculated the `Price vs Avg` column as the percent difference between the car's price and the average for all cars in the dataset. If we want to display the percentages on the values next to the data bars, we can do so by including `percent = TRUE`:

```{r, echo=TRUE}
data <- data %>% mutate("Price vs Avg" = round(Price/mean(Price)-1,2))

reactable(data,
          columns = list(
            `Price vs Avg` = colDef(cell = data_bars_pos_neg(data, percent = TRUE))))
```

## Combining Multiple Formatters 

If we wanted to apply `data_bars()` to one column and `data_bars_pos_neg()` to another, it can be done as shown below:

```{r, echo=TRUE}
reactable(data,
          defaultColDef = colDef(style = list(fontFamily = "monospace", whiteSpace = "pre")), # apply monospace font across all columns
          columns = list(
            Price = colDef(align = "left", # align column header to the left
                           cell = data_bars(data)),
            `Price vs Avg` = colDef(align = "center", # align column header to the center
                                    cell = data_bars_pos_neg(data, percent = TRUE))))
```


