---
title: "2019 NBA Player Ratings"
output: html_document 
resource_files:
  - '.'
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

```{r raptor_table, echo = FALSE}
library(reactablefmtr)
library(htmltools)
library(crosstalk) # for control filters
library(govdown) # used for unstrapping bootstrap theme from crosstalk

### load data from 538
data <- read.csv(file = "RAPTOR_by_team_19-20.csv")

### create shared dataset for crosstalk
crosstalk_data <- SharedData$new(data)

### crosstalk team filter
team_filter <- filter_select(
  id = "team",
  label = "TEAM",
  sharedData = crosstalk_data,
  group = ~ TEAM_NAME
)

### crosstalk conference filter
conference_filter <- filter_select(
  id = "conf",
  label = "CONFERENCE",
  sharedData = crosstalk_data,
  group = ~ CONFERENCE
)

### crosstalk minutes filter
minutes_filter <- filter_slider(
  id = "minutes",
  label = "MINUTES PLAYED",
  sharedData = crosstalk_data,
  column = ~ MIN,
  ticks = TRUE,
  dragRange = FALSE,
  step = 100
)

### load font from google fonts
htmltools::tags$link(href = "https://fonts.googleapis.com/css?family=Chivo:400,600,700&display=swap", rel = "stylesheet")

### match 538's table theme
theme_538 <- function() {
  reactable::reactableTheme(
    searchInputStyle = list(width = "31%", backgroundColor = "#F9F9F9"),
    style = list(
      fontFamily = "Chivo"
    ),
    headerStyle = list(
      "&:hover[aria-sort]" = list(
        background = "hsl(0, 0%, 80%)"),
      "&[aria-sort='ascending'], &[aria-sort='descending']" = list(
        background = "#555",
        color = "#FFF"
      ),
      borderColor = "#333"
    ),
    borderColor = "#CDCDCD"
  )
}

nba_table <-  reactable(
  crosstalk_data,
  theme = theme_538,
  ### add column group header
  columnGroups = list(
    colGroup(name = "OVERALL RAPTOR", columns = c("OFF","DEF","TOT"))
    ),
  showSortIcon = FALSE,
  searchable = TRUE,
  language = reactableLang(
    searchPlaceholder = "SEARCH FOR A PLAYER..."),
  defaultPageSize = 100,
  columns = list(
    TEAM_NAME = colDef(show = FALSE),
    CONFERENCE = colDef(show = FALSE),
    RANK = colDef(maxWidth = 45, name = ""),
    PLAYER = colDef(maxWidth = 225),
    ### add logos using embed_img()
    TEAM_LOGO = colDef(
      name = "TEAM",
      maxWidth = 70,
      align = "center",
      cell = embed_img(height = "25", width = "40")
    ),
    ### add icons using icon_assign()
    MIN = colDef(
      maxWidth = 85,
      align = "center",
      cell = icon_assign(
        data,
        icon = "stopwatch",
        fill_color = "#555",
        buckets = 5
      ),
      style = list(borderRight = "1px dashed rgba(0, 0, 0, 0.3)")
    ),
    ### add color scales using color_scales()
    OFF = colDef(
      maxWidth = 60,
      cell = function(x)
        sprintf("%+0.1f", x),
      style = color_scales(data, colors = c("#fd84a9", "white", "#42c2ca"))
    ),
    ### add color scales using color_scales()
    DEF = colDef(
      maxWidth = 60,
      cell = function(x)
        sprintf("%+0.1f", x),
      style = color_scales(data, colors = c("#fd84a9", "white", "#42c2ca"))
    ),
    ### add color scales using color_scales()
    TOT = colDef(
      maxWidth = 60,
      cell = function(x)
        sprintf("%+0.1f", x),
      style = color_scales(data, colors = c("#fd84a9", "white", "#42c2ca"))
    ),
    ### add bars using data_bars_pos_neg()
    WAR = colDef(
      maxWidth = 280,
      align = "center",
      cell = data_bars_pos_neg(
        data,
        colors = c("#fd84a9", "#fee6ed", "#d9f2f4", "#42c2ca"),
        number_fmt = scales::number_format(accuracy = 0.1)
      ),
      style = list(borderLeft = "1px dashed rgba(0, 0, 0, 0.3)")
    )
  )
)

### display crosstalk filters
### unstrap was used to prevent the crosstalk filters from applying their own bootsrap theme on this site. It is generally not needed.
div(unstrap(bscols(
  widths = c(4, 4, 4),
    unstrap(team_filter),
    unstrap(conference_filter),
    unstrap(minutes_filter)))
)

### display table
div(nba_table)
```

Source: [FiveThirtyEight](https://projects.fivethirtyeight.com/nba-player-ratings/)

Raw Data: [modern_RAPTOR_by_team.csv](https://github.com/fivethirtyeight/data/blob/master/nba-raptor/modern_RAPTOR_by_team.csv)

Cleaned Data: [`RAPTOR_by_team_19-20.csv`](RAPTOR_by_team_19-20.csv) 

### {reactablefmtr} elements used

- `embed_img()` is used to display team logos in the TEAM column. Logo URL's were imported to the cleaned dataset from [{teamcolors}](https://github.com/beanumber/teamcolors)
- `icon_assign()` is used to bucket the number of minutes played into five groups and display as stopwatch icons in the MIN column
- `color_scales()` is used to conditionally color the cells in the three columns under OVERALL RAPTOR
- `data_bars_pos_neg()` is used to display the data bars in the WAR column

### Code

```{r ref.label="raptor_table", eval=FALSE}
```

