---
title: "Getting Started with Sparklines"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(dplyr)
library(palmerpenguins)
```

## Sparklines

In order to use the sparkline functions within {reactablefmtr}, one must first download the [{dataui}](https://timelyportfolio.github.io/dataui/index.html) package from GitHub.

Once installed, users of {reactablefmtr} will have the ability to create highly customizable interactive sparkline line charts and bar charts.

We will start off with an example of the sparkline line charts using data from the [{palmerpenguins}](https://allisonhorst.github.io/palmerpenguins/index.html) package.

The first thing we need to do is summarize the column in a list format:

```{r}
df <- penguins %>%
  filter(!is.na(sex)) %>%
  group_by(species, sex) %>%
  summarize(flipper_length = list(flipper_length_mm))
```

Then we can call `react_sparkline()` within the cell of {reactable}:

```{r}
reactable(df,
          columns = list(
            species = colDef(maxWidth = 85),
            sex = colDef(maxWidth = 85),
            flipper_length = colDef(
              cell = react_sparkline(df))
          ))
```

By default, `react_sparkline()` is interactive and displays the value when we hover over them, but there is an option to turn this off by setting `tooltip` to FALSE.

## Line Options

We can change the color of the line with `line_color`: 

```{r}
reactable(df,
          columns = list(
            species = colDef(maxWidth = 85),
            sex = colDef(maxWidth = 85),
            flipper_length = colDef(cell = react_sparkline(df,
                                                           line_color = "blue"))
          ))
```

If we want to assign line colors to specific groups, we can do so by creating a columns with the color assignments and calling that column name within `line_color_ref`:

```{r}
df <- df %>%
  mutate(
    cols = case_when(
      species == "Adelie" ~ "#f5a24b",
      species == "Chinstrap" ~ "#af52d5",
      species == "Gentoo" ~ "#4c9b9b",
      TRUE ~ "grey"
    )
  )

reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkline(df,
                                                   line_color_ref = "cols"))
  )
)
```

By default, the curvature of the line is of type "cardinal", but we have the option to change it to "linear" as well as "monotoneX", or "basis" within `line_curve`. We can also adjust the width of the line with `line_width`:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        line_width = 2,
        line_curve = "linear",
        line_color_ref = "cols"
      )
    )
  )
)
```

## Area Charts

By setting `show_area` to TRUE, we can show the filled area beneath the line.

Note that we can also increase the height of the sparklines by providing a value within the `height` option to better display the trends within the data. 

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkline(
      df,
      height = 80,
      show_area = TRUE,
      line_color_ref = "cols"
    ))
  )
)
```

The color of the filled area is 90% transparent, but we are able to darken the colors by increasing the opacity within `area_opacity`: 

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        height = 80,
        show_area = TRUE,
        area_opacity = 1,
        line_color_ref = "cols"
      )
    )
  )
)
```

By default, the color of the filled area is assigned from the color of the line, but we can change the color of the area with `area_color`. Alternatively, we can also custom assign colors using `area_color_ref` using the same method we used with `line_color_ref` earlier.

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        height = 80,
        show_area = TRUE,
        area_color = "grey",
        line_width = 3,
        line_color_ref = "cols"
      )
    )
  )
)
```


## Points and Labels

If we wanted to add points to particular data points on the sparkline, we could do so using `highlight_points`. Within `highlight_points`, we can call a helper function, which is also called `highlight_points`, and assign colors to either the min, max, first, last, or all data points.

Below, we are assigning the color red to the minimum values on the sparkline and the color blue to the maximum values: 

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkline(
      df,
      highlight_points = highlight_points(min = "red", max = "blue")
    ))
  )
)
```

We may also apply the labels directly to the sparkline by specifying which values we would like to display with `labels`. The label options are the same as `highlight_points` where we can label either the first, last, min, max, or all values. Note that the `labels` option will work with or without the `highlight_points` option:

```{r}
reactable(df,
          columns = list(
            species = colDef(maxWidth = 85),
            sex = colDef(maxWidth = 85),
            cols = colDef(show = FALSE),
            flipper_length = colDef(cell = react_sparkline(
              df,
              labels = c("first", "last"),
              highlight_points = highlight_points(first = "green", last = "purple")
            ))
          ))
```


## Stat Lines

We may want to display summary statistics about each sparkline series and can do this by using the `statline` option. The statistical summary options that are available are mean, median, min, or max. 

The example below adds a mean reference line to each of the sparklines and displays the mean value to the right of each line:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      minWidth = 200,
      cell = react_sparkline(
        df,
        height = 80,
        line_color = "blue",
        statline = "mean"
      )
    )
  )
)
```

There are additional options to control the appearance of the statline and statline label as well:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        height = 80,
        line_color = "blue",
        statline_color = "orange",
        statline_label_color = "orange",
        statline_label_size = "1.1em",
        statline = "mean"
      )
    )
  )
)
```


## Band Lines

To add a band line to each of the sparklines, we can use the `bandline` option as shown below. The options within `bandline` are "innerquartiles" which shows the inner-quartile range of each series, and "range" which will show the full range of the data from the minimum value to the maximum value.

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        height = 80,
        line_width = 1,
        line_color_ref = "cols",
        bandline = "innerquartiles"
      )
    )
  )
)
```

The color and opacity of the bandline can also be adjusted as shown below:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        height = 80,
        line_color_ref = "cols",
        bandline = "innerquartiles",
        bandline_color = "green",
        bandline_opacity = 0.4
      )
    )
  )
)
```

We may also stack multiple elements together, such as showing the bandline with a mean statline:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkline(
        df,
        height = 80,
        line_color_ref = "cols",
        highlight_points = highlight_points(min = "red", max = "blue"),
        labels = c("min", "max"),
        statline = "mean",
        bandline = "innerquartiles"
      )
    )
  )
)
```


## Bar Charts

To display the sparkline chart as a bar chart rather than a line chart, we can use `react_sparkbar()`:

```{r}
reactable(df,
          columns = list(
            species = colDef(maxWidth = 85),
            sex = colDef(maxWidth = 85),
            cols = colDef(show = FALSE),
            flipper_length = colDef(cell = react_sparkbar(df))
          ))
```

## Bar Options

Many of the options that are available within `react_sparkline()` are also available within `react_sparkbar()` with some few minor differences. For example, if we wanted to assign custom colors to each of the bars, we could use `fill_color_ref`:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkbar(df,
                                                  fill_color_ref = "cols"))
  )
)
```

By default, the line color around each of the bars is transparent, but we can also assign custom colors to the outlines with `line_color_ref`:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(
      cell = react_sparkbar(
        df,
        fill_color = "transparent",
        outline_width = 2,
        outline_color_ref = "cols"
      )
    )
  )
)
```

## Bars and Labels

Another difference in `react_sparkbar` is if we want to highlight particular data points, we would use `highlight_bars` instead of `highlight_points`. The options in which data points to highlight are the same (first, last, min, max, or all).

Note: the height of the bars auto-starts at the minimum value in each series. Therefore, if we assign a color to the minimum value within `highlight_bars`, we will be unable to see it unless we declare the minimum value as a number less than the minimum value present in the dataset:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkbar(
      df,
      height = 80,
      min_value = 160,
      fill_color = "lightgrey",
      highlight_bars =  highlight_bars(min = "red", max = "blue")
    ))
  )
)
```

We can also assign labels using the same method as `react_sparkline()`:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkbar(
      df,
      height = 80,
      fill_color = "lightgrey",
      labels = c("first", "last"),
      label_size = "1em",
      highlight_bars = highlight_bars(first = "green", last = "purple")
    ))
  )
)
```

## Statlines and Bandlines

Just like with `react_sparkline()`, statlines and bandlines can be layered onto `react_sparkbar` using the same options:

```{r}
reactable(
  df,
  columns = list(
    species = colDef(maxWidth = 85),
    sex = colDef(maxWidth = 85),
    cols = colDef(show = FALSE),
    flipper_length = colDef(cell = react_sparkbar(
      df,
      height = 80,
      fill_color_ref = "cols",
      bandline = "innerquartiles",
      statline = "mean"
    ))
  )
)
```

