---
title: "Highlight Min & Max Values"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(reactablefmtr)
library(dplyr)
library(palmerpenguins)
```

## Example

If we wanted to highlight the minimum, maximum, or even both the minimum and maximum values in a particular column, we can do this using `highlight_max()`, `highlight_min()`, or `highlight_min_max()`.

```{r}
data <- sample_n(penguins,10) %>% # sample 10 rows
  filter(!is.na(bill_length_mm)) %>% # remove NA's
  select(species, bill_length_mm, bill_depth_mm, flipper_length_mm)

reactable(data)
```

By default, the `highlight_max()` function identifies the maximum value in a column with green bold text...

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = highlight_max(data))))
```

And the `highlight_min()` function identifies the minimum value in a column with red bold text:

```{r}
reactable(data,
          columns = list(
            bill_depth_mm = colDef(style = highlight_min(data))))
```

If we want to identify both the minimum and maximum value in a column, we can use  the `highlight_min_max()`:

```{r}
reactable(data,
          columns = list(
            flipper_length_mm = colDef(style = highlight_min_max(data))))
```

<br>

## Custom Colors

If we want to change the color of any of either the `highlight_min()` or `highlight_max()` functions above, we can do so by assigning the color with the `font_color` option.

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = highlight_max(data,
                                                          font_color = "blue"))))
```

For `highlight_min_max()`, the minimum value font color can be changed with `min_font_color` and the maximum value font color can be changed with `max_font_color`.

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = highlight_min_max(data,
                                                              min_font_color = "purple", 
                                                              max_font_color = "orange"))))
```

<br>

## Highlight Cells

Additionally, we have the option to highlight the entire cell, rather than just the value, using the `highlighter` option below:

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = highlight_max(data, 
                                                       font_color = "black", 
                                                       highlighter = "gold")),
            bill_depth_mm = colDef(style = highlight_max(data, 
                                                       font_color = "black", 
                                                       highlighter = "gold")),
            flipper_length_mm = colDef(style = highlight_max(data, 
                                                       font_color = "black", 
                                                       highlighter = "gold"))))
```

If we want to do the same within `highlight_min_max`, we just follow a step similar to assigning the min and max font colors with `min_highlighter` and `max_highlighter`:

```{r}
reactable(data,
          columns = list(
            bill_length_mm = colDef(style = highlight_min_max(data, 
                                                             min_font_color = "orange", 
                                                             max_font_color = "purple",
                                                             min_highlighter = "purple", 
                                                             max_highlighter = "orange")),
            bill_depth_mm = colDef(style = highlight_min_max(data, 
                                                             min_font_color = "orange", 
                                                             max_font_color = "purple",
                                                             min_highlighter = "purple", 
                                                             max_highlighter = "orange")),
            flipper_length_mm = colDef(style = highlight_min_max(data, 
                                                             min_font_color = "orange", 
                                                             max_font_color = "purple",
                                                             min_highlighter = "purple", 
                                                             max_highlighter = "orange"))))
```
